---
title: "lab_01"
author: "sean mussenden"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the slice of PPP loans for West Virginia.

**Task** Run the codeblock below to load the data.

```{r}
wv_ppp_loans <- read_rds("lab_01.rds")
```

## Examine Data

**Task** Run the codeblock below to use glimpse to get a sense of the column names, types and sample data.
**Q1** How many rows and columns does this dataframe have?
**A1** [33,336 rows and 58 columns]

```{r}
glimpse(wv_ppp_loans)
```

## Answer questions

Use the pre-lab and your reference notebook as examples.

**Q2** Which West Virginia county got the most loans? How many?
**A2** [Kanawha, 4273]

```{r}
wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n()
  ) %>%
  arrange(desc(count_loans))

```
**Q3** Which West Virginia county got the fewest loans? How many?
**A3** [Palm Beach, 2]

```{r}
wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n()
  ) %>%
  arrange(count_loans)

```
**Q4** What's your best educated guess to explain why the county identified in question 2 had a lot of loans and why the county identified in question 3 had only a few loans? Hint: you won't find the answer in the data, you'll need to do some quick web research. You do not need to write any code for this question, so no codeblock is provided.  
**A4** [My guess is due to both counties population size. For instance, the county with the most loans was Kanahawa, which also happened to be the most populous county in West Virigina, so this seems logical. It makes sense for the most loans to come out of the county with the most people. On the flip side, I couldn't even find a population count for Palm Beach, but I can only assume it was even smaller than some of the counties I found with the smallest population, including Wirt and Doddridge, which had some of the fewest loans as well.]


**Q5** What West Virginia business got the state's largest PPP loan? How much was it for? What does the business do? How many jobs did the loan retain, according to the data?
**A5** [Greenbriar Hotel Corporation, $8912202, it's a hotel/resort company, and it retained 500 jobs]

```{r}
wv_ppp_loans %>%
  arrange(desc(amount))
```

**Q6** What West Virginia politician (with his family) owns the company identified in question 5? What's your source for that information? Hint: this will require web research, you won't be able to answer this purely from the data set.  You do not need to write any code for this question, so no codeblock is provided.  
**A6** [Jim Justice. I found that on Wikipedia, which I then corroborated with another news article that also stated he was the owner. Source: https://wvmetronews.com/2021/07/25/in-swirl-of-financial-turmoil-greenbrier-hotel-corp-is-hit-with-million-dollar-irs-lien/]

**Q7**. Produce a summary table that has the following information for West Virginia as a whole (the whole state, not by county):
* count of loans
* total amount of all loans
* the average loan amount
* the median loan amount
* the smallest loan amount
* the largest loan amount

Store it as a variable object called `wv_summary` so that you can examine it with the data viewer. What newsworthy nuggets jump out at you that you could imagine dropping into a news story? Write a few summary sentences with overall findings.

**A7** [A couple things stand out from the data. One thing I thought was interesting was that Ritchie, which only had 119 loans total had a company within the mark for the highest total loan amount. One single company's loan was making up around 40% of the counties total loan amount. Another thing to look at would be Palm Beach. They only have two loans, both upwards of $600000, making their average much, much higher than any other county. Would be interesting to look at what these two companies are.Perhaps another thing to look at is the situation in Kawahana. The county with by far the most total loans, there are a number of counties with companies that took much higher individual loans. It seems like a number of the biggest companies aren't in the most populous county. By itself that's probably not that important, but when I was looking at the population graphs for individual WV counties, Kawahana has been on a steady decline for the past decade. Perhaps theres a story to written about that possible correlation.]
```{r}
wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n(),
    total_loans_amount = sum(amount),
    mean_loan_amount = mean(amount),
    median_loan_amount = median(amount),
    min_loan_amount = min(amount),
    max_loan_amount = max(amount)
  ) %>%
  arrange(desc(max_loan_amount))
```
